---
title: "Gumbell Distribution"
author: "Kyra Baranick"
format: docx
editor: visual
---

library(dataRetrieval) library(tidyverse)

install.packages("extRemes") library(extRemes)

# Parameter code for discharge = 00060

```{r}
pCode <- "00060"

siteNumber <- "12363000" #Flathead in C Falls 
parameterCd <- "00060" # USGS code for discharge (cfs)
```

# Retrieve daily mean discharge (discharge is always in cfs)

```{r}
daily <- readNWISdv(siteNumber, parameterCd, startDate = "1900-01-01", endDate = Sys.Date())

daily <- renameNWISColumns(daily)
```

# Oct 1–Sep 30); or use calendar year if preferred

```{r}
daily <- daily %>% mutate(waterYear = ifelse(month(Date) >= 10, year(Date) + 1, year(Date)))
```

# Compute annual maxima

```{r}
ams <- daily %>% group_by(waterYear) %>% summarize(annual_max = max(Flow, na.rm = TRUE), .groups = "drop")
```

# View

```{r}
head(ams)

hist(log(ams$annual_max))

acf(log(ams$annual_max))

plot(ams$waterYear,ams$annual_max)

#Max flood is declining!!!

#fit to gumbel 
gumbel_fit_ams<- fevd(ams$annual_max, type = "Gumbel") print(gumbel_fit_ams)

#pdf of gumbel
dg<- function(x, mu, beta){ z<-(x-mu)/beta (1/beta)* exp(-(z+exp(-z))) }

x<-seq(10,150000, by=100)

plot(x, dg(x, mu= 39500, beta=12681))

#cdf of gumbel dg_cdf <- function(x, mu, beta){

exp(-exp(-(x-mu)/beta))

}
```

#### 

#Max flood is declining!!!

#fit to gumbel gumbel_fit_ams\<- fevd(ams\$annual_max, type = "Gumbel") print(gumbel_fit_ams)

#pdf of gumbel dg\<- function(x, mu, beta){ z\<-(x-mu)/beta (1/beta)\* exp(-(z+exp(-z))) }

x\<-seq(10,150000, by=100)

plot(x, dg(x, mu= 39500, beta=12681))

#cdf of gumbel dg_cdf \<- function(x, mu, beta){

exp(-exp(-(x-mu)/beta))

}

#plot pQ plot(x, 1-dg_cdf(x, mu= 39500, beta=12681))

pQ\<- 1-dg_cdf(x, mu= 39500, beta=12681)

dQ\<- 1e-6\*x\^2.5 #made this up

plot(x, dq) #looks good?

pQ_future\<- 1-dg_cdf(x, mu= 50000, beta=12681) #move future gumbel location to 50Kcfs

##plot them up

plot(x, pQ*dQ, ylim=c(0,500000)) points(x, pQ_future*dQ)

#########Clark Fork for Rachel

siteNumber \<- "12340500" \# Clark Fork parameterCd \<- "00060" \# USGS code for discharge (cfs)

\# Retrieve daily mean discharge (discharge is always in cfs) daily \<- readNWISdv(siteNumber, parameterCd, startDate = "1900-01-01", endDate = Sys.Date())

daily \<- renameNWISColumns(daily)

\# Extract water year (Oct 1–Sep 30); or use calendar year if preferred daily \<- daily %\>% mutate(waterYear = ifelse(month(Date) \>= 10, year(Date) + 1, year(Date)))

\# Compute annual maxima ams \<- daily %\>% group_by(waterYear) %\>% summarize(annual_max = max(Flow, na.rm = TRUE), .groups = "drop")

\# View head(ams)

hist(log(ams\$annual_max))

acf(log(ams\$annual_max))

plot(ams$waterYear,ams$annual_max)

write.csv(ams, file="/Users/bob.hall/Dropbox/RachelJ/ams.csv")

\#### #Max flood is declining!!!

#########Dec rain on snow pCode \<- "00060"

siteNumber \<- "12358500" \# MF Flathead parameterCd \<- "00060" \# USGS code for discharge (cfs)

\# Retrieve daily mean discharge (discharge is always in cfs) daily \<- readNWISdv(siteNumber, parameterCd, startDate = "1900-01-01", endDate = Sys.Date())

daily \<- renameNWISColumns(daily)

\# Extract water year (Oct 1–Sep 30); or use calendar year if preferred daily \<- daily %\>% mutate(waterYear = ifelse(month(Date) \>= 10, year(Date) + 1, year(Date)))

daily \<- daily %\>% mutate(waterYear = ifelse(month(Date) \>= 10, year(Date) + 1, year(Date)), WY_start = make_date(waterYear - 1, 10, 1), wy_day = as.numeric(Date - WY_start) + 1)

##WY 2026 mf_2026\<- read.csv("/Users/bob.hall/Dropbox/USGS value of data/MF_2026.csv") mf_2026$Date<- as.Date(mdy_hms(mf_2026$datetime))

mf_2026_daily\<- mf_2026 %\>% group_by(Date) %\>% summarise( Flow=mean(Flow)) mf_2026_daily \<- mf_2026_daily %\>% mutate(waterYear = ifelse(month(Date) \>= 10, year(Date) + 1, year(Date)), WY_start = make_date(waterYear - 1, 10, 1), wy_day = as.numeric(Date - WY_start) + 1)

ggplot(daily, aes(x=wy_day, y=Flow))+ geom_line(color="black", alpha=0.1)+ geom_line(data=mf_2026_daily, aes(x=wy_day, y=Flow), color="red")+ theme_classic()+ ylim(0,30000)
