---
title: "Initial code"
author: "Kyra Baranick"
format: pdf
editor: visual
---

# Load Packages

```{r}
library(dataRetrieval)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(extRemes)
library(terra)
library(sf)
library(writexl)
```

```{r}
stationID <- c(
"USGS-13011000", 
#Snake River NR Morgan WY 
"USGS-13296500",
#SALMON RIVER BL YANKEE FORK NR CLAYTON ID 
"USGS-06188000",
#Lamar River nr Tower Ranger Station YNP 
"USGS-09188500",
#GREEN RIVER AT WARREN BRIDGE, NEAR DANIEL, WY 
"USGS-13336500",
#SELWAY RIVER NR LOWELL ID
 "USGS-12355500",
 # Station Name - N F Flathead River nr Columbia Falls MT (MT) 
 "USGS-12414500",
#ST JOE RIVER AT CALDER ID 
 "USGS-06207500", 
#Station Name: Clarks Fork Yellowstone River nr Belfry MT 
 "USGS-13337000"
#Station name LOCHSA RIVER NR LOWELL ID
)
```

# Pulling metadata for each watershed

```{r}

site_info <- read_waterdata_monitoring_location(monitoring_location_id = c(
"USGS-13011000", 
#Snake River NR Morgan WY 
"USGS-13296500",
#SALMON RIVER BL YANKEE FORK NR CLAYTON ID 
"USGS-06188000",
#Lamar River nr Tower Ranger Station YNP 
"USGS-09188500",
#GREEN RIVER AT WARREN BRIDGE, NEAR DANIEL, WY 
"USGS-13336500",
#SELWAY RIVER NR LOWELL ID
 "USGS-12355500",
 # Station Name - N F Flathead River nr Columbia Falls MT (MT) 
 "USGS-12414500",
#ST JOE RIVER AT CALDER ID 
 "USGS-06207500", 
#Station Name: Clarks Fork Yellowstone River nr Belfry MT 
 "USGS-13337000"
#Station name LOCHSA RIVER NR LOWELL ID
),
  properties = c("monitoring_location_id",
    "site_type",
    "drainage_area",
                                                          "monitoring_location_name"))

knitr::kable(site_info)
```

```{r}
#all avlible itme series metadata (when it started and ended)
ts_available <- read_waterdata_combined_meta(monitoring_location_id = c("USGS-12355500",
"USGS-12340000","USGS-06289000", "USGS-06207500","USGS-13337000"),
                                  parameter_code = c("00060", "00010"))
```

```{r}
#creating map ofinvestigated gage sites
library(leaflet)

leaflet_crs <- "+proj=longlat +datum=WGS84" #default leaflet crs

leaflet(data = site_info |> 
                sf::st_transform(crs = leaflet_crs)) |> 
    addProviderTiles("CartoDB.Positron") |> 
    addCircleMarkers(popup = ~monitoring_location_name, 
                   radius = 0.1,
                   opacity = 1)

```

# Reading in all daily flow data

```{r}
daily_flows <- read_waterdata_daily(
  monitoring_location_id = stationID,
  parameter_code = "00060",      # discharge
  statistic_id = "00003",        # mean
  time = c("1800-10-01", "")
)


```

# Cleaning to only include date, discharge and the site with name mutated

```{r}
daily_clean <- daily_flows %>%
  select(
    site_id = monitoring_location_id,
    date = time,
    discharge_cfs = value
  )

daily_clean <- daily_clean %>%
  mutate(site_name = site_info$monitoring_location_name[
      match(site_id, site_info$monitoring_location_id)
    ],discharge_m3s = discharge_cfs * 0.0283168)
#adding water yera column
daily_clean <- daily_clean %>%
  mutate(
    date = as.Date(date),
    watyear = if_else(
      month(date) >= 10,
      year(date) + 1,
      year(date)
    )
  )



```

# Extract Annual Maxima

```{r}

annual_max <- daily_clean %>%
  group_by(site_id, site_name, watyear) %>%
  slice_max(discharge_cfs, with_ties = FALSE) %>%
  mutate(discharge_m3s = discharge_cfs * 0.0283168) %>%
  ungroup()

```

# Make Gumbel Curve

## Start with the functions

```{r}
gumbel_pdf <- function(x, mu = 0, beta = 1) {
  if (beta <= 0) stop("Scale parameter beta must be > 0")
  z <- (x - mu) / beta
  (1 / beta) * exp(-(z + exp(-z)))
}
gumbel_cdf <- function(x, mu = 0, beta = 1) {
  exp(-exp(-(x - mu) / beta))
}

```

# Make Flood Freq. Curve

```{r}
floodfreq_daily <- daily_clean %>%
  filter(discharge_m3s > 1) %>%        # remove zero/negative flows
  group_by(site_name) %>%
  arrange(desc(discharge_m3s)) %>%
  mutate(
    m = row_number(),
    years = n_distinct(watyear),       # number of water years in record
    P = m / (years + 1),               # exceedance probability
    T = 1 / P                          # return period in YEARS
  ) %>%
  ungroup()


```

## plotting with log scales

```{r}
ggplot(floodfreq_daily, aes(x = T, y = log(discharge_m3s), color = site_name)) +
  scale_x_log10() +
  geom_point(alpha = 0.2, size = 1.0) +
  geom_line(alpha = 0.7) +
  labs(
    x = "Return Period (years)",
    y = "ln(Discharge m³/s)",
    title = "Daily-Value Flood Frequency Curves (Natural Log Scale)"
  ) +
  theme_bw()


```

# Simple Time Series

```{r}
 library(ggplot2)
 
ggplot(
  annual_max, aes(x = discharge_m3s))+
  geom_histogram()+
  facet_wrap (~site_name, scales = "free_y")
)

```

```{r}
ggplot(
  annual_max, aes(x = watyear, y = discharge_m3s))+
  geom_point()+
  facet_wrap (~site_name, scales = "free_y")
```

# Slope of annual snow melt season vs time (this needs work)

```{r}
daily_clean_manipulated <- daily_clean %>%
  mutate(doy = yday(date)) %>%                 # add day of year
  filter(doy >= 0, doy <= 365) %>%            # melt season
  group_by(site_name, watyear) %>%
  arrange(date) %>%
  mutate(dQ = discharge_m3s - lag(discharge_m3s)) %>%  # daily change
  filter(dQ > 0) %>%                            # rising days only
  summarise(rise_slope = mean(dQ, na.rm = TRUE)) %>%
  ungroup()

```

```{r}
ggplot(daily_clean_manipulated, aes(watyear, rise_slope)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~ site_name, scales = "free_y") +
  labs(
    x = "Water Year",
    y = "Mean Rising-Limb Slope (m3s/day)",
    title = "Snowmelt Rising-Limb Slope Trends Across Sites"
  )

```

# Peak dates over time 

```{r}

annual_max <- annual_max %>%
  mutate(
    wy_start = as.Date(paste0(watyear - 1, "-10-01")),
    daywatyear = as.numeric(date - wy_start) + 1
  )

ggplot(annual_max, aes(x = watyear, y = daywatyear)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~ site_name, scales = "free_y") +
  labs(
    title = "Earlier Snowmelt Peak Over Time",
    x = "Water Year",
    y = "Day of Water Year of Peak Flow"
  )

```
