---
title: "Initial code"
author: "Kyra Baranick"
format: pdf
editor: visual
---

# Load Packages

```{r}
library(dataRetrieval)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(extRemes)
library(terra)
library(sf)
library(writexl)
```

```{r}
stationID <- c(
 "USGS-06043500",# Gallatin River (at Gallatin Gatway, MT)
 "USGS-12332000",# Middle Fork Rock Creek (near Philipsburg, MT)
 "USGS-09196500", # Pine Creek (Above Freemont Lake, WY)
 "USGS-06090500", # Belt Creek (near Monarch, MT)
 "USGS-12358500",# Middle Fork Flathead River(Near West Glacier, MT)
 "USGS-06188000", # Lamar River (Near Tower Ranger Station, WY)
 "USGS-09188500",# Green River (At Warren Bridge, near Daniel, WY)
 "USGS-13336500", # Selway River (Near Lowell, ID)
 "USGS-12355500", # North Fork Flathead River (Flathead British Colombia, MT)
 "USGS-06207500", # Clarks Fork Yellowstone River (Near Belfry, MT)
 "USGS-13337000", # Locsha River (Near Lowell, ID)
 "USGS-06078500", # North Fork Sun River (Near Augusta, MT)
 "USGS-06079000", # South Fork Sun River (Near Augusta, MT)
 "USGS-05014500", # Swiftwater Creek (At Many Glacier, MT)
 "USGS-12354000",# St. Regis River (Near St. Regis, MT)
 "USGS-06200000", # Boulder River (at Big Creek, MT)
 "USGS-06102500", #Teton River (Below South Fork Near Choteau, MT) 
 "USGS-12414500" # St. Joe River (at Calder, ID)
)
```

# Pulling metadata for each watershed

```{r}

site_info <- read_waterdata_monitoring_location(monitoring_location_id = c(
 "USGS-06043500",# Gallatin River (at Gallatin Gatway, MT)
 "USGS-12332000",# Middle Fork Rock Creek (near Philipsburg, MT)
 "USGS-09196500", # Pine Creek (Above Freemont Lake, WY)
 "USGS-06090500", # Belt Creek (near Monarch, MT)
 "USGS-12358500",# Middle Fork Flathead River(Near West Glacier, MT)
 "USGS-06188000", # Lamar River (Near Tower Ranger Station, WY)
 "USGS-09188500",# Green River (At Warren Bridge, near Daniel, WY)
 "USGS-13336500", # Selway River (Near Lowell, ID)
 "USGS-12355500", # North Fork Flathead River (Flathead British Colombia, MT)
 "USGS-06207500", # Clarks Fork Yellowstone River (Near Belfry, MT)
 "USGS-13337000", # Locsha River (Near Lowell, ID)
 "USGS-06078500", # North Fork Sun River (Near Augusta, MT)
 "USGS-06079000", # South Fork Sun River (Near Augusta, MT)
 "USGS-05014500", # Swiftwater Creek (At Many Glacier, MT)
 "USGS-12354000",# St. Regis River (Near St. Regis, MT)
 "USGS-06200000", # Boulder River (at Big Creek, MT)
 "USGS-06102500", #Teton River (Below South Fork Near Choteau, MT) 
 "USGS-12414500" # St. Joe River (at Calder, ID)
),
  properties = c("monitoring_location_id",
    "site_type",
    "drainage_area",
                                                          "monitoring_location_name"))

knitr::kable(site_info)
```

# Cleaning site info to just include river names

```{r}
site_info_clean <- site_info |>
  dplyr::mutate(
    site_id = gsub("USGS-", "", monitoring_location_id)
  )

```

```{r}
site_info_clean <- site_info_clean |>
  dplyr::mutate(
    river_name = monitoring_location_name |>
      gsub(",.*", "", x = _) |>                 # remove everything after comma
      gsub(" nr .*", "", x = _, ignore.case=TRUE) |> 
      gsub(" near .*", "", x = _, ignore.case=TRUE) |> 
      gsub(" at .*", "", x = _, ignore.case=TRUE) |> 
      gsub(" above .*", "", x = _, ignore.case=TRUE) |> 
      gsub(" below .*", "", x = _, ignore.case=TRUE) |> 
      toupper()
  ) |>
  dplyr::select(site_id, river_name)

```

```{r}
#creating map ofinvestigated gage sites
library(leaflet)

leaflet_crs <- "+proj=longlat +datum=WGS84" #default leaflet crs

leaflet(data = site_info |> 
                sf::st_transform(crs = leaflet_crs)) |> 
    addProviderTiles("CartoDB.Positron") |> 
    addCircleMarkers(popup = ~monitoring_location_name, 
                   radius = 0.1,
                   opacity = 1)

```

# Reading in Continuous flow data

# Reading in all daily flow data

```{r}
daily_flows <- readRDS("daily_flows.rds")
```

```{r}

inst_flows <- readRDS("inst_flows.rds")
```

# Cleaning to only include date, discharge and the site with name mutated

```{r}

#adding water yera column
daily_clean <- daily_flows %>%
  mutate(
    site_name = river_name,
    date = as.Date(Date),
    watyear = if_else(
      month(date) >= 10,
      year(date) + 1,
      year(date)
    )
  )



```

## Option to exclude some rivers

```{r}
daily_clean <- daily_clean %>%
  filter(!site_id %in% c("06078500", "06079000"))
```

```{r}

inst_clean <- inst_flows %>%
  mutate(
    datetime = as.Date(datetime),
    watyear = if_else(
      month(datetime) >= 10,
      year(datetime) + 1,
      year(datetime)
    )
  )


```

# Extract Annual Maxima for inst

```{r}

annual_max_inst <- inst_clean %>%
  group_by(site_id, river_name, watyear) %>%
  slice_max(discharge_m3s, with_ties = FALSE) %>%
  ungroup()

```

# Extract Annual max for daily

```{r}
annual_max_daily <- daily_clean %>%
  group_by(site_id, site_name, watyear) %>%
  slice_max(discharge_m3s, with_ties = FALSE) %>%
  ungroup()
```

# Make Flood Freq. Curve

```{r}
floodfreq_daily <- daily_clean %>%
  filter(discharge_m3s > 1) %>%        # remove zero/negative flows
  group_by(river_name) %>%
  arrange(desc(discharge_m3s)) %>%
  mutate(
    m = row_number(),
    years = n_distinct(watyear),       # number of water years in record
    P = m / (years + 1),               # exceedance probability
    T = 1 / P                          # return period in YEARS
  ) %>%
  ungroup()


```

## plotting with log scales

```{r}
ggplot(floodfreq_daily, aes(x = T, y = log(discharge_m3s), color = river_name)) +
  scale_x_log10() +
  geom_point(alpha = 0.2, size = 1.0) +
  geom_line(alpha = 0.7) +
  labs(
    x = "Return Period (years)",
    y = "ln(Discharge m³/s)",
    title = "Daily-Value Flood Frequency Curves (Natural Logged y-values)"
  ) +
  theme_bw()


```

# Simple Time Series

```{r}
 library(ggplot2)
 
ggplot(annual_max_inst, aes(x = discharge_m3s)) +
  geom_histogram() +
  facet_wrap(~ river_name, scales = "free") +
  labs(
    x = "Discharge (m³/s)",
    y = "Count",
    title = "Instantaneous Value Annual Max"
  )+
  theme(
    strip.text = element_text(size = 8)  # smaller facet titles
  )


 ggplot(annual_max_daily, aes(x = discharge_m3s)) +
  geom_histogram() +
  facet_wrap(~ river_name, scales = "free") +
  labs(
    x = "Discharge (m³/s)",
    y = "Count",
    title = "Daily Value Annual Max"
  )+
 theme(
    strip.text = element_text(size = 5)  # smaller facet titles
  )

```

```{r}
ggplot(
  annual_max_inst, aes(x = watyear, y = discharge_m3s))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap (~river_name, scales = "free")+
  labs(
    y = "Discharge (m³/s)",
    x = "Water Year",
    title = "Annual Max Discharge time series by Water Year (continuous)"
  )+
  theme(
    strip.text = element_text(size = 5)  # smaller facet titles
  )

ggplot(
  annual_max_daily, aes(x = watyear, y = discharge_m3s))+
  geom_point()+
  geom_smooth(method = lm)+
  facet_wrap (~river_name, scales = "free") +
  labs(
    y = "Discharge (m³/s)",
    x = "Water Year",
   title = "Annual Max Discharge time series by Water Year (daily)"
  )+ 
  theme(
    strip.text = element_text(size = 5)  # smaller facet titles
  )
  
```

# Peak dates over time

```{r}
annual_max_inst <- annual_max_inst %>%
  mutate(
    datetime = as.Date(datetime),
    wy_start = as.Date(paste0(watyear - 1, "-10-01")),
    day_wy = as.numeric(datetime - wy_start) + 1
  )
annual_max_daily <- annual_max_daily %>%
  mutate(
    datetime = as.Date(date),
    wy_start = as.Date(paste0(watyear - 1, "-10-01")),
    day_wy = as.numeric(date - wy_start) + 1
  )

```

```{r}

ggplot(annual_max_inst, aes(x = watyear, y = day_wy)) +
  geom_point() +
geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ river_name, scales = "free_y") +
  labs(
    x = "Water Year",
    y = "Day of water year continaing peak flow (1-365)",
    title = "Instantaneous Value Peak date Over time "
  )+
  theme_minimal()


 ggplot(annual_max_daily, aes(x = watyear, y = day_wy)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ river_name, scales = "free_y") +
  labs(
    x = "Water Year",
    y = "Day of water year continaing peak flow (1-365)",
    title = "Daily Value Peak Date"
  )+
   theme_minimal()

```

# Center of mass

```{r}

library(fasstr)
library(dplyr)

```

```{r}
river_lookup <- daily_clean %>%
  distinct(site_id, river_name)


centofmass <- daily_clean %>%
rename(Date = Date,
       STATION_NUMBER = site_id, 
       Value = discharge_m3s)

flow_timing_forCOM <- calc_annual_flow_timing(centofmass) %>%
  left_join(
    river_lookup,
    by = c("STATION_NUMBER" = "site_id")
  )



  

```

```{r}
library(ggplot2)
ggplot(flow_timing_forCOM, aes(x = Year , y = DoY_50pct_TotalQ)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  facet_wrap(~river_name, scales = "free") +
  labs(
    title = "Center of Mass for 50% of total Q",
    x = "Water Year",
    y = "Day of Water Year of 50% of Total Annual Flow"
  )
```

# Annual Low Flows

```{r}
daily_clean <- daily_clean %>%
  mutate(
    wateryear = ifelse(month(Date) >= 10,
                       year(Date) + 1,
                       year(Date)),
    doy = as.numeric(Date - as.Date(paste0(wateryear - 1, "-10-01"))) + 1
  )

annual_low <- daily_clean %>%
  group_by(site_id, watyear) %>%
  slice_min(discharge_m3s, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(site_id, river_name, watyear, Date, discharge_m3s, doy)


```

```{r}
library(ggplot2)

ggplot(annual_low,
       aes(x = watyear, y = doy)) +
  geom_point(color = "steelblue") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  facet_wrap(~ river_name) +
  labs(
    title = "Timing of Annual Minimum Flow",
    x = "Water Year",
    y = "Day of Water Year (Minimum Flow)"
  ) +
  theme_bw()


```
